\include{6001mac}
\begin{document}
\psetheader{IAP 2005}{Lecture 5}

\section*{Problems}

\begin{enumerate}
\item Write {\tt list-copy}, which takes a list and returns an
identical new list (ie do not just return the original list, {\tt
cons} up a new list).
\begin{verbatim}
(list-copy (list 1 2 3))
;Value: (1 2 3)
\end{verbatim}
\vspace{1in}

\item Write {\tt n-copies}, which takes a value and a number of
copies, and returns a list with the appropriate number of copies.
\begin{verbatim}
(n-copies 7 5)
;Value: (7 7 7 7 7)
(n-copies "yay" 1)
;Value: ("yay")
(n-copies 7 0)
;Value: ()     ; or #f
(n-copies (list 3) 3)
;Value: ((3) (3) (3))
\end{verbatim}
\vspace{1in}

\item Write {\tt reverse}, which takes a list and returns new list
with the order of the elements reversed.
\begin{verbatim}
(reverse (list 1 2 3))
;Value: (3 2 1)
(reverse (list 1))
;Value: (1)
\end{verbatim}
\vspace{1.5in}

\item Write {\tt append}, which takes two lists and returns a new list
with the elements of the first list and the second list.
\begin{verbatim}
(append (list 3 4) (list 1 2))
;Value: (3 4 1 2)
(append nil (list 1 2))
;Value: (1 2)
\end{verbatim}
\vspace{1in}

\item Write {\tt list-ref}, which takes a list and an index (starting
at 0), and returns the nth element of the list.  You may assume that
the index is less than the length of the list.

\begin{verbatim}
(list-ref (list 17 42 35 "hike") 0)
;Value: 17
(list-ref (list 17 42 35 "hike") 1)
;Value: 35
(list-ref (list 17 42 35 "hike") 2)
;Value: 35
\end{verbatim}
\vspace{1in}

\item Write {\tt list-range}, which takes two numbers (a,b : a < b)
and returns a list containing the numbers from a to b, inclusive.

\begin{verbatim}
(list-range 1 5)
;Value: (1 2 3 4 5)
(list-range 2 5)
;Value: (2 3 4 5)
(list-range 42 42)
;Value: (42)
(list-range 207 5)
;Value: ()
\end{verbatim}
\vspace{1in}

\item Write {\tt max-list}, which takes in a list of numbers and
returns the maximum element.  You may assume that the list is
non-empty.  (Hint: different base case than normal!)
\begin{verbatim}
(max-list (list 1))
;Value: 1
(max-list (list 1 3 5))
;Value: 5
(max-list (list 2 56 8 43 21))
;Value: 56
\end{verbatim}
\vspace{1in}
\end{enumerate}

\section*{Data Abstraction}

\begin{enumerate}
\item Derived Type - A user-designated and implemented type.
\item Constructor - Builds entity of the type
\item Selector - Returns one of the values of the type
\item Contract - Specifies the relationship between the constructor(s)
and the selector(s).
\end{enumerate}

\begin{verbatim}
(define (make-point x y)

(define (get-x point)

(define (get-y point)

\end{verbatim}

\begin{enumerate}
\setcounter{enumi}{7}
\item Write {\tt add-points} which takes two points and returns a new
point which is the sum of the x and y coordinates.
\begin{verbatim}
(define result (add-points (make-point 3 4) (make-point 1 2)))
(get-x result)
;Value: 4
(get-y result)
;Value: 6
\end{verbatim}
\vspace{1in}

\item Write {\tt left-of?} which takes two points and returns true if
the first point is to the left of the second point.

\begin{verbatim}
(left-of? (make-point 3 4) (make-point 1 2))
;Value: #f
(left-of? (make-point -3 4 (make-point 1 2)))
;Value: #t
\end{verbatim}
\vspace{1in}
\end{enumerate}

\subsection*{Stacking Abstractions: Segments}

\begin{enumerate}
\setcounter{enumi}{9}
\item Implement an abstraction for line-segments, which are defined by
a pair of end-points.
\vspace{2in}

\item Write {\tt segment-length}, which takes a segment and returns
it's length.
\vspace{1in}
\end{enumerate}

\end{document}
