\input 6001mac.tex
\def\fbox#1{%
  \vtop{\vbox{\hrule%
              \hbox{\vrule\kern3pt%
                    \vtop{\vbox{\kern3pt#1}\kern3pt}%
                    \kern3pt\vrule}}%
        \hrule}}

\def\emptybox#1#2{\fbox{\vbox to #2{\hbox to #1 {\null}}}}

\begin{document}

\psetheader{IAP, 2005}{Quiz I}

\centerline{\bf Closed Book -- one sheet of notes}

\vskip 10pt

Separately, we have distributed an answer sheet.  You may use the
space on the exam booklet for whatever temporary work you find useful,
but you {\bf MUST} enter your answers into the answer sheet.  {\bf
Only the answer sheet will be graded.}  Each problem that requires an answer has
been numbered.  Place your answer at the corresponding number in the
answer sheet.

Note that any procedures or code fragments that you write will be
judged not only on correct function, but also on clarity and good
programming practice.

The answer sheet asks for your section number and tutor.  For your
reference, here is the table of section numbers.

\vskip 0.25in

\begin{center}
\begin{tabular}{|c|c|c||c|c|}
\hline
Section & Time &        Room &  Instructor &    Tutor \\ \hline
R01 & MTWRF 10 AM &   32-144 & Ben Vandiver & Ben Vandiver \\ \hline
\end{tabular}
\end{center}

\newpage
{\bf Part 1: (18 points)}

For each of the following expressions you are to (1) draw the box and pointer
diagram corresponding to the list or pair structure created, and (2) write the
printed result of evaluating the last expression in the sequence.

\medskip

{\small \begin{verbatim}
(define a (list 1 (cons 2 nil) 3))
a
\end{verbatim}}
\vspace*{-0.1in}
\question{Draw box and pointer for {\tt a}}
\question{Printed result}

\rule{4in}{.5pt}
\medskip
{\small \begin{verbatim}
(define x (list 1 2))
(define y (list 3 (list 4)))
(define z (append x y))
z
\end{verbatim}}
\vspace*{-0.1in}
\question{Draw box and pointer for {\tt z}}
\question{Printed result}

\rule{4in}{.5pt}
\medskip
{\small \begin{verbatim}
(define y (list 42 17 54))
(define z (cons 7 (cdr y)))
z
\end{verbatim}}
\vspace*{-0.1in}
\question{Draw box and pointer for {\tt z}}
\question{Printed result}


\newpage
{\bf Part 2: (20 points)}

We will consider the following function:
$\mathrm{tower}(x,n)=x^{(x^{(x^{...})})}$ (where $x$ appears $n$ times).
For example, $\mathrm{tower}(2,3)=2^{(2^{(2^{1})})}=16$.  By definition
$\mathrm{tower}(x,0)=1$.  You should use {\tt (expt x n)} which
computes $x^n$ to help implement {\tt tower} below.

\begin{verbatim}
(define (tower x n)
  INSERT1)
\end{verbatim}

\question{What expression should be used for INSERT1?}


\newpage
{\bf Part 3: (30 points)}

Ben Bitdiddle (a common character in 6.001 with no connection to your
instructor) is implementing a rating tracking system for people
playing the card game Spades.  This tracking system maintains a count
of wins and losses for each player.  His first job is to implement a
player abstraction that wraps up the player's name, number of wins,
and number of losses.  He's managed to implement the selectors, but
the constructor continues to elude him.  Help Ben Bitdiddle out by
completing the constructor in such a way that the contract is
preserved (ie the selectors return the appropriate values).

\begin{verbatim}
(define (make-player name wins losses)
  INSERT1)

(define (player-name player)
  (first player))

(define (player-wins player)
  (second player))

(define (player-losses player)
  (third player))


; example usage:
(define p (make-player "Ben" 5 3))

(player-name p)
;Value: "Ben"
(player-wins p)
;Value: 5
(player-losses p)
;Value: 3
\end{verbatim}

\question{What expression should be used for INSERT1?}

{\bf -----------ABSTRACTION BARRIER--------------}\\

The number of primary interest to the players (and Ben) is their win
ratio, which is the ratio of wins to total games.

\begin{verbatim}
(define (player-win-ratio player)
  INSERT2)
\end{verbatim}

\question{What expression should be used for INSERT2?}

Finally, given a list of players who have only played each other, for
every win counted by one player there should be a loss counted by
another.  Thus, the total number of wins must equal the total number
of losses.  We can use this fact to check to see if anyone is cheating
and misreporting their won/loss record.

Ben is writing a procedure called {\tt check-records} which takes a
list of players and returns 0 if the number of wins equals the number
of losses, returns a positive number if there are more wins than
losses, and a negative number if there are more losses than wins.
Help him finish the procedure.

\begin{verbatim}
(define (check-record list-of-players)
  (if (null? list-of-players)
      INSERT3
      (+ INSERT4
         (check-record (cdr list-of-players)))))
\end{verbatim}

\question{What expression should be used for INSERT3?}

\question{What expression should be used for INSERT4?}

\newpage
{\bf Part 4:  (12 points)}

Indicate whether the following procedures are iterative or recursive.

\begin{verbatim}
(define (slow-mul x y)
  (if (= x 0)
      0
      (+ y (slow-mul (- x 1) y))))

(define (slow-add x y)
  (if (= x 0)
      y
     (slow-add (- x 1) (+ y 1))))

(define (watch-this n)
  (if (= n 0)
      1
      (if (< n 0)
          (watch-this (- n))
          (* n (watch-this (- n 1))))))

(define (excitement)
  (excitement))
\end{verbatim}

\question{Is {\tt slow-mul} iterative or recursive?}
\question{Is {\tt slow-add} iterative or recursive?}
\question{Is {\tt watch-this} iterative or recursive?}
\question{Is {\tt excitement} iterative or recursive?}

\newpage
{\bf Part 5:  (20 points)}

For each of the following expressions or sequences of expressions,
state the value returned as the result of evaluating the final
expression in each set, or indicate that the evaluation results in an
error.  If the expression does not result in an error, also state the
``type'' of the returned value, using the notation from lecture. If
the result is an error, state in general terms what kind of error
(e.g. you might write ``error: wrong type of argument to procedure'').
If the evaluation returns a built-in procedure, write {\tt primitive
procedure}, and its {\tt type}. If the evaluation returns a
user-created procedure, write {\tt compound procedure}, and also
indicate its {\tt type} using the notation we introduced in class.
You may assume that evaluation of each sequence takes place in a newly
initialized Scheme system.

For example, for the expression {\tt 88} your answer would be \\
\begin{tabular}{|l|l|}\hline
Value: & 88\\\hline
Type: & number\\\hline
\end{tabular}

\medskip
\vspace*{-0.1in}
{\small \begin{verbatim}
((lambda (a b) (a b))
 (lambda (c) (* 2 c))
 10)
\end{verbatim}}
\question{Value}
\question{Type}
% 20, number

\bigskip
\vspace*{-0.1in}
{\small \begin{verbatim}
(lambda (m n)
  (if m (* 2 n) (/ 2 n)))
\end{verbatim}}
\question{Value}
\question{Type}
% compound procedure, (boolean, number) -> number

\bigskip
\vspace*{-0.1in}
{\small \begin{verbatim} 
(let ((x <)
      (y 10)
      (z 20))
  (z y x))
\end{verbatim}}
\question{Value}
\question{Type}
% error, 20 is not applicable

\bigskip
\vspace*{-0.1in}
{\small \begin{verbatim}
(define (double x) (* 2 x))
(define (check y)
  (if (< (double y) 10)
      "yip"
      (if (> (double y) 6)
          "yay"
          "yuck")))
check
\end{verbatim}}
\question{Value}
\question{Type}
%compound procedure, anytype -> number

\bigskip
\vspace*{-0.1in}
{\small \begin{verbatim}
(define three 
  (lambda (a b c) (* a b c)))
(three (three 1) (three 2) (three 3))
\end{verbatim}}
\question{Value}
\question{Type}
%% error, wrong number arguments to ``three''

\end{document}

